---
title: "Presentation of Netflix Title Analysis"
author: "Hudson Colletti"
subtitle: "May 8, 2025"
format:
  revealjs:
    scrollable: true
    slide-number: true
    show-slide-number: all
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

## Overview of Data

-   Analysis of Netflix Movies & TV Shows\
-   Data from [TidyTuesday (2021-04-20)](https://github.com/rfordatascience/tidytuesday)
-   Focus on trends in U.S. content, title length, and naming conventions\
-   Data originally collected by [Shivam Bansal on Kaggle](https://www.kaggle.com/datasets/shivamb/netflix-shows/data)

------------------------------------------------------------------------

## Visualization 1 - Netflix Content Trends Over Time in the United States

```{r}
#| echo: false
library(tidyverse)
library(tidytuesdayR)
library(ggplot2)
library(stringr)

tuesdata <- tidytuesdayR::tt_load('2021-04-20')
netflix <- tuesdata$netflix
netflix_summary <- netflix |>
  mutate(added_year = str_extract(date_added, "\\d{4}")) |>
  mutate(united_states_only = str_extract(country, "^United States")) |>
  mutate(title_length = str_length(title)) |>
  filter(str_detect(added_year, "\\d{4}")) |>
  filter(str_detect(country, "^United States")) |>
  group_by(added_year, type) |>
  summarize(count = n(), .groups = "drop")

ggplot(netflix_summary, aes(x = added_year, y = count, group = type, color = type)) +
  geom_line() +
  geom_point() +
  labs(title = "Netflix Content Trends Over Time in the United States",
       x = "Year Added to Netflix",
       y = "Number of Titles",
       color = "Content Type")

```

## Visualization 1 Analysis

From 2015 onward, there is a steep increase in content added to Netflix in the U.S. Movies drive most of the growth, but TV Shows also rise sharply.

-   The sudden drop in 2021 likely reflects:
    -   COVID-19 production slowdowns
    -   Partial year data
    -   Internal Netflix policy shifts

Increased User Retention

## Visualization 2 - Average Netflix Title Length Over Time

```{r}
#| echo: false
title_length_by_year_type <- netflix |>
  group_by(release_year, type) |>
  filter(!str_detect(title, "\\bThe\\b")) |>
  summarize(avg_title_length = mean(str_length(title)), na.rm = TRUE, count = n(), .groups = "drop") |>
  filter(release_year >= 2000)

ggplot(title_length_by_year_type, aes(x = release_year, y = avg_title_length, color = type)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Average Netflix Title Length Over Time",
    x = "Release Year",
    y = "Average Title Length (characters)",
    color = "Content Type"
  )
```

## Visualization 2 Analysis

-   Movie title lengths have fluctuated over time, with consistently longer titles after 2015.
-   TV show title lengths fluctuated from 2000-2012
    -   Since 2012, TV titles have stabilized between 15–17 characters
-   Movie titles are now typically longer than TV Shows

## Visualization 3 - Netflix Titles Starting with 'A' or 'An' Over Time

```{r}
#| echo: false
netflix |>
  mutate(added_year = str_extract(date_added, "\\d{4}")) |>
  filter(str_detect(added_year, "\\d{4}")) |>
  group_by(added_year) |>
  summarize(
    total_titles = n(),
    titles_with_article = sum(str_detect(title, "^A\\b|^An\\b")),
    proportion = titles_with_article / total_titles,
    .groups = "drop"
  ) |>
  ggplot(aes(x = added_year, y = proportion)) +
  geom_point() +
  labs(
    title = "Proportion of Netflix Titles Starting with 'A' or 'An' Over Time",
    x = "Year Added to Netflix",
    y = "Proportion of Titles"
  )

```

## Visualization 3 Analysis

-   Between 2008-2014 the proportion of titles beginning with 'A'/'An' was extremely low. 2011 and 2013 are outlier years possibly explained by:
    -   Low total title counts
    -   Certain acquisitions that skewed overall
-   The proportion of titles beginning with these articles increased sharply from 2015 on
    -   This growth may reflect a gaming of alphabetization in recommendation interfaces.
-   These proportions stabilize between 1.5-2.5% of all titles.

## Summarized Conclusions

-   Netflix’s title trends reflect strategic business choices:

    -   Rapid content expansion
    -   Title naming decisions

-   These patterns provide insight into how data can shift business models in the streaming industry

-   As the overall number of titles released on Netflix grew, the proportion of titles with 'A'/'An' increased as well.

# Thanks for Watching

Any Questions?
